A Software Engineering Perspective on \\ OpenMP and PThreads

### 2017-08-07 sections/01-introduction.tex
Add text to responses section, improve figures

@@

In this context, the approach of using Compilation Directives to guide the
compiler in the parallelization process has gained popularity. These directives
are implemented using the compiler's pre-processing directives and are utilized
as annotations that provide tips about the sequential code. Among the tools and
extensions that use compilation directives are the \texttt{OpenMP} API
\cite{Dagum1998a} \cite{Chapman:2007} used for writing programs for multi-core
architectures, and the \texttt{OpenACC} programming
standard~\cite{openacc:api}, used for writing programs for heterogeneous
CPU/GPU architectures. The \texttt{OpenMP 4.0} specification supports
offloading capabilities \cite{openmp:api:2013} such as \texttt{OpenACC} and
\texttt{AMD HSA}~\cite{amd:hsa:site}. These tools aim to make easier the task
of designing and implementing efficient and provably correct parallel
algorithms. However, the abstraction layers built by the extensions also have
the potential to conceal architectural details and to generate new parallel
programming challenges.

When comparing \texttt{OpenMP} to other available tools for developing
multi-threaded applications such as the \texttt{pthreads} interface, the API
can be considered almost costless. However, when writing more complex
applications this may not be true. The design and implementation of such
complex applications require more knowledge about platforms and execution
models than what can be shown by tutorials, that present high level and trivial
examples. These tutorials try to convince the user that programming with
directives is simple and easy to do, which is not always true. For instance, to
achieve good performance in some cases it is necessary to describe many
restrictions on annotations~\cite{OpenMPTasks2009}~\cite{mattson2003good}.

Krawezik and Cappello~\cite{CPE:CPE905} evidenced this difficulty by comparing
MPI and three OpenMP versions of benchmarks. The most popular use of
\texttt{OpenMP} directives is at the loop level. The programmer discovers
potentially parallel loops and annotates the code with directives starting and
closing parallel regions. Good results were obtained only when SPMD was
implemented on OpenMP, and achieving good performance required a significant
programming effort~\cite{CPE:CPE905}.

In this work we have gathered data from assignments made by graduate students
for the \emph{Introduction to Parallel and Distributed Computing} course. Among
other exercises, the students were asked to find examples of \texttt{OpenMP}
code in tutorials and official manuals that failed under special conditions.
These conditions could be created by, for example, placing barriers or forcing
racing conditions by exposing inaccurate thread management.

The rest of the paper is organized as follows. The following section discusses
related work...

>>>>

\todo[inline,color=cyan,author=Pedro]{Write new introduction}

@@

In this work we have gathered data from assignments made by graduate students
for the \emph{Introduction to Parallel and Distributed Computing} course. Among
other exercises, the students were asked to ...

\todo[inline,color=cyan,author=Pedro]{Describe paper structure}

### 2017-08-07 sections/02-background.tex

The approaches of the last paragraph introduce tools for parallel programming
before the related concepts.  Instead, some learning approaches
consider learning software tools as just one of the objectives of the course,
and argue that their introduction should come after the introduction of
concepts of Parallel and Distributed Computing.

>>>>

The approaches mentioned so far introduce tools for parallel programming before
the related concepts.  Other approaches instead consider learning software
tools as just one of the objectives of the course, and argue that their
introduction should come after the introduction of concepts of Parallel and
Distributed Computing.
